Does the leaf economic spectrum hold within plant functional types? A Bayesian multivariate trait meta-analysis

### 2019-09-06 ecoapp-comments/response.txt

        337 I don’t see how you arrive at this conclusion. If the data-generating process was multivariate normal, univariate and multivariate means should be identical. To me, this result rather suggests that the correlation is not multivariate, which seems to question the assumptions of your model, or that the PFT means are not in line with the global mean (so that you get a mixture distribution, but this has nothing to do with multivariate). It may be a good idea to provide at least a few visual checks of model adequacy, e.g. by posterior predictive model checks

        339 OK, I think if the traits are correlated, the multivariate model is clearly better, because we want to know this correlation. I find this so obvious that it wouldn’t even need discussing. However, with the comments above in mind, I think you should specify what you mean by “bias” – as said, I don’t think there should be a bias in the mean if the data-generating process is multivariate normal, and you fit univariate normal distributions.

  The univariate and multivariate means are only identical if partially missing data are dropped, in which case additional information about missing parameters based on covariance with existing parameters is neglected. However, by leveraging among-trait covariance, we can make more educated guesses about parts of the trait space that have data for other traits, and we can therefore at least partially overcome sampling biases. We have added a demonstration of this effect using simulated data as Method S2 in the Supplementary Information, and reference it here (line 563).

        341 What do you mean by “reliable”?

  We changed this to say "too few observations to estimate covariance patterns for some PFTs with much statistical power" (line 567).

        349 That makes sense to me, but I was missing the relating info in the results

  See our above responses (comment "As a side note...", between "233..." and "244...") about the power analysis and pairwise sample size tables in the Supplementary Information.

        355 It also seems obvious that the procedure could make use of phylogenetic or morphological similarity, as it is commonly done in PGLS type models

  We have mentioned this as an additional future direction of this work (line 584-585).

        356 It seems what you discuss here is independent of PFTs, right? But then I would say people are already working on this, and it doesn’t really relate to the results of this study

  We respectfully disagree with this comment for several reasons. A key result of this work is the additional constraint on trait estimates provided through multivariate approaches, whether hierarchical or not, and we expect this constraint to be particularly valuable for the hard-to-measure traits discussed in this section. Moreover, if, as we show (more clearly in this revision), even well-established leaf economic relationships have some hierarchical character, then examining the extent to which this is the case for other "economic spectra" (namely, wood and roots, as mentioned here) is a very relevant future application.

        368 I think the mutually confounding issue is actually the point that is most interesting from an ecological perspective. The relevant question is if trade-offs such as the LES scale through all levels (individual, PFTs, all plants), or if they could be different at some scales. When taking together your results and other literature that looks at LES within species, it seems to me that current evidence suggestions that trade-offs tend to be universal? Could you extend on this discussion, possibly also highlighting which trade-offs you suspect to be universal, and which might be different at different ecological scales?

  As mentioned in the main comments, our revised version of the manuscript focuses much more on the differing extent to which some trait relationships are universal or actually scale-dependent.

        372 I wouldn’t sign this statement, but OK, if you think so

        385 See my concern: uncertainty != predictive error

  We have removed this section, so these comments are no longer relevant.

        Fig. 1 not a big fan of this lattice plot layouts, you are losing a lot of space with the plot headings, and the PFTs would better be placed below the plots, under the bars, imo. Consider if the visualization can be improved.

        Fig. 2 as said, not sure what you mean by significant. Moreover, why would you not plot the n.s. too? A correlation of zero is also an information, isn’t it?

        Fig. 3 as said in the previous comments, I wonder if this could be better condensed. Difficult to see any patterns because of the high amount of info.

  We have replaced all three of these figures in the revised version.

        Fig. 4 Same point about lattice layout. Also, not sure why you suddenly introduce the units to the traits, while they are absent in all other plots.

  The units in this figure make more sense now, since it is the first figure to actually show trait values (rather than their slopes). We have decided to keep the "lattice" layout because it is the most effective way we have to show results for all 8 traits. However, we have moved the results from the univariate and multivariate models to the supplementary information to reduce visual clutter and improve readability.

        Fig. 5 see my main comment: CI is not fit.

  See main response.

2 Reviewer 2
============

        This manuscript starts from the relatively simple question of whether leaf economic spectrum (LES) patterns (correlations between SLA and leaf N, etc) hold when they are applied to a subset of the global distribution of plants (divided into PFTs). This is an interesting question because we use the LES a lot in trait-driven land surface models, yet these relationships tend to break down using smaller subsets (by area, etc) of plants. Reaching beyond the traits traditionally considered in the LES, this paper also considers metabolic 'traits'.

        While I think this is an interesting question, I found myself wondering at the end of this paper what the actual point was. It reads a bit like a bunch of exploratory analyses in search of a conclusion, as evidenced by the many nearly unreadable multi-panel and matrix plots. Given that we know quite a bit about the LES, it would have been interesting to start this manuscript with some hypotheses instead of an exploration -> looking at all of these plots it is difficult to untangle what is unexpected versus what is normal. Similarly, why did the authors think the hierarchical model was necessary, and what does it mean that it didn't improve estimates? Overall this manuscript feels a bit too focused on the technical methods at the expense of the motivating questions.

  We appreciate this point and agree that the introduction did not do a good enough job of laying out the motivation for this study, and that its driving hypotheses were implicit, not explicit, making things more difficult for the reader. We have revised the introduction to be clearer and more logically organized, with an explicit motivating question at its end. (See main response.)

        Writing-wise, this manuscript is well written and clear, however, it feels like it is written to target people already in the small niche of ecological land surface modelers. I think it would be more appealing to a wider audience if the introduction was edited to explain more of the underlying concepts and why we care about them. For example, the manuscript refers to LES working at the global scale and not at local scales, then refers to PFTs as an intermediate scale (lines 74-76) but this is using "scale" in a very abstract way that deserves more explanation. Global and local scales are both spatial, but PFTs are really a conceptual scale between lumping everything together ('global') and modeling each species or individual separately.

  We have re-written large parts of the introduction to address this concern, with particular emphasis on the need for PFTs in the context of trait ecology and vegetation modeling.

        At the very end of the manuscript PFTs are then semi-equated with biomes (line 383-384) which isn't right.

  We have removed this paragraph.

        Throughout the introduction it would help if the manuscript spent a bit more time explaining to a novice/outside reader why PFTs, traits, and land surface models are all worth reading and learning about.

  See earlier comment about revisions to the introduction.

        Table 1: Make sure it's clear that the top 8 PFTs in the list are trees.

  We have merged this table with the sample size table (see next comment), and have replaced this accordingly.

        Figure 1: Seems like this would be better as a table in an appendix? Dedicating a whole figure to show sample sizes seems like a lot.

  We have replaced this figure with a table.

        Figure 2: This figure needs more explanation in the caption of what is being shown. What does the slope and length of the lines represent? Also, it's almost impossible to tell this many colors apart in a figure (colorbrewer doesn't even have 14 classes as an option) - it might be worth trying to come up with a pared down version of this that focuses on some key points.

        Figure 3: Is this showing something substantially different from figure 2? This one is even harder to interpret just given the sizes of the plots and the numbers of bars.

  We have replaced both of these figures with a single, very different figure (Fig. 1) that, we hope, does a better job of highlighting the key results.

        Figure 4: Again, this is a lot of information making it hard to see what is important and what isn't, especially given that in most cases the three model types produce almost identical results. Also, the caption states that leaf lifespan and SLA only show mass based results, but is there a mass vs. area difference for leaf lifespan? and SLA is per mass by definition, no? Finally on this one, it looks like CLM values have confidence interval bars on them, but they are just single values, I think?

  We have moved the original version of this figure to the supplementary information and replaced it with a version that only has the results from the hierarchical model and the CLM 4.5 parameters. We have also changed the symbol for CLM 4.5 values from "+" to "x" to make it clearer that those values do not have error bars. Finally, we have clarified in the caption that the normalization refers to traits /other/ than SLA or leaf lifespan.

        Tables S1 and S2: SLA is listed as kg m-2 in both of these - should be m2 kg-1 (as in text, line 90), and/or flipped to LMA for area-normalized?

  Thanks; yes, these units should have been m2 kg-1. We have fixed this in the revision.

        Table S4: What do 'present' and 'missing' mean here? or how are there correlation values when 'present' = 0?

  We have replaced this table, so this comment no longer applies.
